Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications

### 2014-12-12 report/paper.tex
review 2 paper

This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the additive subgroup

\[
L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
\] of $\mathbb{R}^m$, a lattice with basis $b_1,b_2,...,b_n$.
A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors will be noted in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.

@@

A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$. The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $B'$ by performing a combination of the following column operations:
\begin{itemize}
\item Swap two columns
\item Negate the elements of a column
\item Addition of an integer multiple of one column to another
\end{itemize}

@@

For full dimensional lattices, the volume can be computed:
\[
\text{vol} \,L_{\text{full}} = |\det B|
\] where $B$ is a matrix of the linearly independent lattice's basis vectors.

@@

Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$. Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$. Then $C$ contains at least a nonzero point of the lattice $L$.

A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero vector $x$ such that

\[
\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}
\]

This gives us an upper-bound on the size of a shortest vector in a lattice.

@@

The \emph{Hermite Normal Form} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique. It is the analogue of the reduced echelon form for integer matrices.

A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when\cite{SchnorrStanfordNotes} :
\begin{enumerate}
\item $a_{ij} = 0$ for $j > i$, i.e. $A$ is lower triangular. 
\item $a_{ii} > 0$ for $i=1,2,...,n$
\item $0 \leq a_{ij} < a_{ii}$ for $j < i$
\end{enumerate}

\subsection{Special Lattices}

\subsubsection{Random Lattice}

@@

Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
$$\begin{array}{ccccc}
P & \vec{a_2} &  & \\
\vec{a_2} & 1&   & \\ 
\dots & &  & \\ 
\vec{a_{m}} & & &1 
\end{array}$$
where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.

@@

We define a \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a $n \times n$ square matrix whose entries are $k$-bits picked at random. These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.

@@

The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}. The orthogonality defect can be interpreted geometrically as the product of the basis vector lengths divided by the volume of the lattice. As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller. The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$. If $\text{OrthDefect} = 1$, the basis is completely orthogonal.

@@

The GSO process  can be defined recursively:
\[
\begin{cases}
\vec{b_1}^{\star} = \vec{b_i}  \\
\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b_j^\star}
\end{cases}
\]
where $\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?

@@

\subsection{$\delta$-LLL Reduced}
We say that a basis is $\delta$-LLL reduced if:
\begin{itemize}
\item $| \mu_{i,j}| \leq \frac{1}{2} \text{, } \forall i > j$
\item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
\end{itemize}

@@

Since the LLL-reduction algorithm returns a $\delta$-LLL reduced basis, we can observe the following properties on the basis:
\begin{itemize}
\item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
\item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
\item $\| b_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
\end{itemize} 
where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$.

@@

The LLL algorithm, like many reduction algorithms, can be broken down into two parts: size-reduction and vector swapping. Size-reduction reduces the vectors from the input basis. After that, the vectors of the basis are checked against the Lovász condition. If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ don't satisfy the condition, we swap them and go back to the size reduction. my{Procedure 1..}

@@

Lagrange was the first to formalize\cite{lagrangeArithmetique} the notion of reduction for rank-two lattices.

@@

Lagrange's algorithm solves the lattice reduction problem exactly for lattice bases of dimension 2. It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers. GCD can be thought to be a one dimension reduction problem. The GCD is the first minimum of the lattice $n \mathbb{Z} + m \mathbb{Z}$ spanned by $n$ and $m$.

@@

\end{algorithmic}
\end{algorithm}
\subsection{The Generic Jacobi Method}
We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice. This method was introduced~my{\cite{originalJacobiMethodLatticeBasisReduction}} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.

From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis such as:
\begin{align}
\langle \vec{b_i} \vec{b_j} \rangle &\leq \frac{1}{2} \| \vec{b_i}\|^2 \text{ (for all $1 \leq i < j \leq n$)}.\\
\| \vec{b_i}\| &\leq \|\vec{b_j} \| \text{ (for all $1 \leq i < j \leq n$)}
\end{align}

@@

\subsubsection{$\omega$-Lagrange-reduced conditions}
An $\omega$ is introduced to 
There are two conditions for a basis to be $\omega$-Lagrange-reduced.

@@

\end{cases}
\]  where $1/\sqrt{3} \leq \omega < 1$.

@@

\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $}

@@

\FOR{every column $k$ of $G$}
\STATE{$g_{lk} = g_{kl} $}
\ENDFOR

@@

\end{algorithmic}
\end{algorithm}
\my{Please make sure that there is no mistake above. I will check it later. Syntax}

@@

We implemented this Jacobi algorithm in C++ with the \emph{newNTL} library. The FPLLL implementation was used for LLL reductions.

@@

Our experimentation focused on the quality of the reduction. The two quality metics we used for comparing the reductions are:

@@

\subsection{Input bases}
Numerous papers were written about the Jacobi method for lattice reduction by S. Qiao \cite{originalJacobiMethodLatticeBasisReduction}\cite{complexityAnalysisOfJacobiMethod}\cite{enhancedJacobi} but none of the papers are clarifying what bitsize was used for the randomly generated entries of the vectors. The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance. MIMO uses smaller dimensions lattices. \cite{enhancedJacobi} and \cite{tian2014hybrid}  are using lattices that are at most 20 and  100 dimensions respectively. These dimensions are inappropriate for cryptographic applications.

In our experiments, we have used a bitsize of around 10 bits entries when generating the almost orthogonal bases.

@@

We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.

\begin{figure}[h!]
  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
  \centering
    \includegraphics[width=0.5\textwidth]{random-matrix-omega-06defect.png}
\end{figure}

\begin{figure}[h!]
  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
  \centering
    \includegraphics[width=0.5\textwidth]{random-matrix-omega-09defect.png}
\end{figure}

These orthogonality defects are similar to that the authors found in \cite{tian2014hybrid}.

@@

In the first case, LLL will stop while Jacobi does exchange $B-2$ and $b_3$. In the second case the Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
\end{enumerate}

@@

\section{Conclusions}

@@

\section{Credits}

>>>>

This can be translated into a more formal definition~\cite{SchnorrStanfordNotes}.\newline 
Let $\vec{b_1}, \vec{b_2}, ... \vec{b_n} \in \mathbb{R}^m$ be linearly independent vectors for $n\leq m$. We call the \my{discrete} additive subgroup

\[
L(\vec{b_1}, \vec{b_2},..., \vec{b_n}):= \displaystyle\sum_{i=1}^{n} \vec{b_i} \mathbb{Z}
\] of $\mathbb{R}^m$, a lattice with basis \my{ bold..}$b_1,b_2,...,b_n$.
A lattice that has $\vec{b_1},\vec{b_2},...,\vec{b_n}$ as basis vectors \my{is fully described by this basis which we denote} in matrix form as $B$ where each vector $\vec{b_i}$ is a column of the matrix.
The \emph{dimension of the lattice} is $dim(L) := rank(L) := n$.

@@

A \emph{unimodular matrix} $U$ is a square integer matrix such that $det(U) = \pm 1$. The right multiplication of a matrix $B$ with a unimodular matrix $U$ yields $\my{B'}$ by performing a combination of the following column operations:
\begin{itemize}
\item Swap two columns
\item Negate the elements of a column
\item Addition of an integer multiple of one column to another
\end{itemize}

@@

For full dimensional lattices, the volume can be computed:
\[
\text{vol} \,L_{\text{full}} = |\det B|
\] where $B$ is a matrix of the linearly independent lattice basis vectors.

@@

\my{grammar} Minkowski's Convex Body Theorem tells us that for a full-rank lattice $L$ of $\mathbb{R}^n$. Let $C$ be a measurable subset of $\mathbb{R}^n$, convex, symmetric with respect to 0, and of measure $> 2^n \text{vol}(L)$. Then $C$ contains at least a nonzero point of the lattice $L$.

A corollary of Minkowski's theorem tells us that there exists for any $n$-dimensional full-rank lattice $L$ a nonzero \my{lattice-}vector $x$ \my{whose length is bounded by $\sqrt{n} (\det L)^{1/n}$. So we can derive a bound on the length of a shortest lattice-vector:}

\[
\lambda_1(L) \leq \sqrt{n} (\det L)^{1/n}.
\]

@@

The \emph{Hermite Normal Form}~\cite{SchnorrStanfordNotes} representation of a basis matrix $B$, abbreviated \emph{HNF (B)}, is unique. It is the analogue of the reduced echelon form for integer matrices.

A full-rank matrix $[a_{ij}] \in M_{n,n} (\mathbb{R})$ is in HNF when:
\begin{enumerate}
\item $a_{ij} = 0$ for $j > i$, i.e. $A$ is lower triangular. 
\item $a_{ii} > 0$ for $i=1,2,...,n$
\item $0 \leq a_{ij} < a_{ii}$ for $j < i$
\end{enumerate}

\subsection{Special Lattices \my{Two types of lattices}}
\my{For our experiments, we will take into account two types of lattices: Random lattices as used in cryptology and special lattices that allow an almost orthogonal basis.}
\subsubsection{Random Lattice}

@@

Without loss of generality, we hence restrict tests to random lattices of volume $P$ whose basis in HNF form is as follows:
\my{$$\begin{array}{ccccc}
P & \vec{a_2} &  & \\
\vec{a_2} & 1&   & \\ 
\dots & &  & \\ 
\vec{a_{m}} & & &1 
\end{array}$$}
where $a_i \in \mathbb{Z}/ P\mathbb{Z}$.

@@

We define a\my{n} \emph{almost orthogonal lattice basis} $M$ of dimension $n$ and of bit length $k$ as a\my{n} $n \times n$ square matrix whose entries are $k$-bit \my{integers} picked at random. These lattices are frequently used for signal processing applications~\cite{originalJacobiMethodLatticeBasisReduction}.

@@

The orthogonality defect is a quality indicator that we will use to compare the reduction of \emph{the set of basis vectors}. The orthogonality defect can be interpreted geometrically \my{you recite the definition, no geometrical interpretation is given??} as the product of the basis vector lengths divided by the volume of the lattice. As the basis vectors are getting increasingly orthogonal during a reduction, the nominator will become smaller. The lower bound of the orthogonality defect is $\text{OrthDefect} \ge 1$. \my{In the special case that ..}If $\text{OrthDefect} = 1$, the basis is completely orthogonal.

@@

The GSO process  can be defined recursively:
\[
\begin{cases}
\vec{b_1}^{\star} = \vec{b_i}  \\
\vec{b_i}^{\star} = \displaystyle\sum^{i-1}_{j=1} \mu_{i,j} \vec{b_j^\star}
\end{cases}
\]
where \my{$\mu_{ij} = ..$. Is the set of vectors $b_i^*$ a basis or not? In which case yes?}

@@

\subsection{$\delta$-LLL Reduced}
We say that a basis is $\delta$-LLL reduced if:
\begin{itemize}
\item $| \mu_{i,j}| \leq \frac{1}{2} $ \my{better: for $1 \leq j < i \leq	n$}$\text{, } \forall i > j$
\item Lovász's condition: $\forall (\vec{b_i, b_{i+1}})$, we have $(\delta - \mu^2_{i+1,i}) \|\vec{b}^{\star}_{i}\|^2 \leq \| \vec{b}^{\star}_{i+1} \|^2$
\end{itemize}

@@

Since the LLL-reduction algorithm returns a $\delta$-LLL reduced basis, we can observe the following properties on the basis:
\begin{itemize}
\item $\| \vec{b_1} \| \leq \alpha^{(n-1)/2} \lambda_1 $
\item $\max_i \| \vec{b_i} \| \leq \alpha^{(n-1)/2} \lambda_n $
\item $\| \vec{b}_1 \| \leq \alpha^{(n-1)/4} \det B^{1/n}$
\end{itemize} 
where the parameter $\alpha = 1/(\delta - \frac{1}{4})$. Since $\delta$ is bounded by $1/4 < \delta < 1$, parameter $\alpha \geq 4/3$. \my{actually $>$ }

@@

The LLL algorithm, like many reduction algorithms, can be broken down into two parts: size-reduction and vector swapping. Size-reduction reduces the vectors from the input basis. After that, the vectors of the basis are checked against the Lovász condition. If a pair of vectors of the basis $(\vec{b_{i}}, \vec{b_{i+1}})$ \my{don't} satisfy the condition, we swap them and go back to the size reduction. \my{link to Procedure 1..}

@@

Lagrange was the first to formalize the notion of reduction for rank-two lattices~\cite{lagrangeArithmetique}.

@@

Lagrange's algorithm solves the lattice reduction problem \my{there is no such thing as a lattice reduction problem; you mean it solves the SVP exactly} exactly for lattice bases of dimension 2. It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers. GCD can be thought to be a one dimension reduction problem. The GCD is the first minimum of the lattice $n \mathbb{Z} + m \mathbb{Z}$ spanned by $n$ and $m$.

@@

\end{algorithmic}
\end{algorithm}
\subsection{The Generic Jacobi Method}
We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice. This method was introduced~\cite{originalJacobiMethodLatticeBasisReduction} by Qiao in 2012 claiming experimental results to be considerably faster than LLL.

From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis \my{replace 'such as' by 'for which'}:
\begin{align}
\langle \vec{b_i} \vec{b_j} \rangle &\leq \frac{1}{2} \| \vec{b_i}\|^2 \text{ (for all $1 \leq i < j \leq n$)}.\\
\| \vec{b_i}\| &\leq \|\vec{b_j} \| \text{ (for all $1 \leq i < j \leq n$)}
\end{align}

@@

\subsubsection{$\omega$-Lagrange-reduced conditions}
An \my{A parameter} $\omega$ is introduced to \my{...? }
There are two conditions for a basis to be $\omega$-Lagrange-reduced.

@@

\end{cases}
\]  where $1/\sqrt{3} \leq \omega < 1$ \my{and where $g_{ij}$}.. .

@@

\STATE{$q \leftarrow \lfloor \frac {g_{ll}}{g_{ss}} \rceil $ \my{?? sure? }}

@@

\FOR{every column $k$ of $G$ \my{onlu for the ones smaller than $\ell$? columns or rows?}}
\STATE{$g_{lk} = g_{kl} $}
\ENDFOR

@@

@@

We implemented the \my{variant on the} Jacobi reduction in C++ \my{by use of} with the \emph{newNTL} library. The FPLLL implementation \my{of LLL} was used \my{to compare the results after reduction.}

@@

Our experimentation focused on the quality of the reduction. The two quality \my{metics} we used for comparing the reductions are:

@@

\subsection{Input bases}
Numerous papers were written about the Jacobi method for lattice reduction by S. Qiao \cite{originalJacobiMethodLatticeBasisReduction}\cite{complexityAnalysisOfJacobiMethod}\cite{enhancedJacobi} but none of the papers are clarifying what bitsize was used for the randomly generated entries of the vectors. The main focus of the authors was the use of lattice reduction for signal processing, more specifically multiinput and multioutput (MIMO), the use of multiple antennas at both the transmitter and receiver to improve communication reliance. MIMO uses smaller dimensions lattices. \my{In} \cite{enhancedJacobi} and \cite{tian2014hybrid} \my{a reduction is performed on} lattices of dimensions 20 to 100. These dimensions are inappropriate for cryptographic applications.

In our experiments, we have used a bitsize of around 10 bits entries when \my{actually always} generating the almost orthogonal bases.

@@

We initially ran our Jacobi reduction variant on almost orthogonal bases to attempt to reproduce Qiao's results.
The orthogonality defects are similar to \my{what} the authors found in \cite{tian2014hybrid}. \my{Our results are shown in Fig.~\ref{fig1} and Fig.~\ref{fig2}. We vary $\omega$..... We observe that...}

\begin{figure}[H]
\begin{minipage}{.5\textwidth}

  \centering
    \includegraphics[width=0.9\textwidth]{random-matrix-omega-06defect.png}
      \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.6$}
  \label{fig1}
\end{minipage}
\begin{minipage}{.5\textwidth}

  \centering
    \includegraphics[width=0.9\textwidth]{random-matrix-omega-09defect.png}

  \caption{LLL with delta $\delta = 0.99$ and Jacobi with $\omega = 0.99$}
  \label{fig2}
\end{minipage}

\end{figure}

@@

In the first case, LLL will stop while Jacobi does exchange $\vec{b}_2$ and $\vec{b}_3$. In the second case the Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
\end{enumerate}

@@

\section{Conclusions}
\my{Short paragraph}
\section{Credits \my{ or Acknowledgment}}

@@
